---
title: "Portfolio exam - Part 1 | Methods 1 F23, CogSci @AU"
author: "Vilma Bejer Kristiansen, Benjamin Tang Hørning Jensen, Katarina Krajcova, Zuzanna Katarzyna Zyla and Alexander Juby-Rasmussen GROUP 4"
date: "12/10/2023"
output: html_document:
  toc: true
---

---

#### Deadline: October 12 2023 h23:59

#----------------------------------
# GENERAL INSTRUCTIONS
#----------------------------------

# 1. GITHUB:

# When you are done completing the below code, please push to Github with a commit comment (use _ for spaces in the message)

# You can push as many times as you would like. The last version that you commit before the assignment deadline is the one that will be reviewed for feedback.

# Please use your individual Github account to push assignments. Do not create a group account if you are working in a group.

#----------------------------------

# 2. PUSH CODE & KNIT HTML MARKDOWN DOCUMENT.

# Please push your code AND a knit HTML file of the document to Github. 

#----------------------------------

# 3. GROUPWORK

**Please write your name in the `author` field above, followed by your Group #.**

# If you are working in groups, it should be possible to grade your code AND your written text individually. 

# Therefore, please indicate your group number at the top of your markdown file and indicate your group members' names below.

#----------------------------------

# 4. Suppressing code output.

# Please suppress code output that is not for evaluation (e.g., printouts of data frames, variable values, etc.) by setting echo = FALSE in chunks where only code is being run. (Look in Class3A_PHB_Rmd from class 3 for more info on echo = FALSe and also further down in this document). This will make it much easier to evaluate your HTML output files.

#----------------------------------

#### Group members' names.

# Please indicate your group members' names here.
My group members' names are: []

#### Introduction

The goal of this exam is to write a short data mining report on the CogSci Intro Week Personality Test Data in which you answer the following questions **in prose, code and graphs**.

First of all, let's start by looking at the setup chunk. If you need to load packages or set your working directory, do so here:

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)
## your code goes here
pacman::p_load(tidyverse,pastecs)
```

Now you have to import the personality data from the 'cogsci_personality_2023.csv' file; i.e., the one we have worked with in both class 2 and class 3. Note the filename is in lowercase and it is located in the /renv_cogsci_e23 folder on UCloud. 

Once you have imported it, use the `head()` function to print the first 10 lines of the data set.

```{r load_data}
## your code goes here
setwd('/work/renv_cogsci_e23')
data <- read_csv('cogsci_personality_2023.csv')
head(data)
```

Once you are done loading the data, you can start working on the questions below.

### Question 1

#### Question 1.1

Who can hold their breath the longest on average — those with right or left ocular dominance? Notice that the column is called `ocular_dom`, and that right ocular dominance is indicated in the column with 'Right', while left ocular dominance is indicated in the column with 'Left'. Therefore, you want to only filter out the data in this column which corresponds to either "Right" or "Left".

Plot the data using `ggplot2` to find out the answer to your question. The plots should include error bars depicting the standard error of the mean: you can add these using the `geom_errorbar()` function and specifying `stat = "summary", fun.data = "mean_se"`. 

#### Question 1.2

Then use the `mean()` and `sd()` functions *within a tidyverse pipe* to make a summary data set, in which you show mean and standard deviation of the two eye dominance groups.

Bonus question: If you feel brave, you can instead try making a boxplot (`geom_boxplot()`) or a violin plot (`geom_violin()`) which are better at representing the actual distribution of the data (compared to a bar plot, which only depicts mean and standard deviation).


```{r question1}
## your code goes here
## Vilma
#1.1 Solution:

filter_data <- data %>%
  filter(ocular_dom != "Both") #Filter the data, in the Ocular_dom collum, so we don't have "Both" in the dataset.

filter_data %>% ggplot(aes(x=ocular_dom, y=breathhold, fill = ocular_dom)) + #making the bar plot, with the variables.
  geom_bar(
       stat = "summary",
    fun = mean
    
  ) + # Adding error bars, for the x/y values. 
  
  geom_errorbar(
    stat = "summary",
    fun.data = mean_se
    
    
  ) + #Cleaning up the look, so it is nicer to look at
  
  labs(
    x= "Ocular Dominance",
    y= "Breathold (s)",
    title= "Breathold according to Ocular Dominance"
    
    
  ) +
  
  
  theme_minimal()


```


```{r}
## Vilma
#1.2 Solution:

data %>%
 group_by(ocular_dom) %>%
  summarise(mean(breathhold),
            sd(breathhold)) #Grouping the data only to look at the Ocular_dom's mean and SD related to the numeric value from "Breathhold" 



```

Explain your results in plain terms here (max 3 sentences): 

###Vilma: The right ocular dominance sample, can hold their breath the longest on average (Right_mean:56.65, left_mean:55.44), but the one with the longest breathold is in the left sample group (SD_left:30.55, SD_right:19.71). There is no error-bars on the "both", because there is only one who answered both, so no SD.

---

#### Question 2

Who likes silence vs. noise best -- by gender? Also in this case you should plot the data using `ggplot2` (including error bars depicting the standard error of the mean), then use the `mean()` and `sd()` functions to find mean and standard deviation of the two genders (still making a summary data set with tidyverse and pipes).

Bonus question: If you feel brave, you can instead try making a boxplot (`geom_boxplot()`) or a violin plot (`geom_violin()`) which are better at representing the actual distribution of the data (compared to a bar plot, which only depicts mean and standard deviation).

```{r question2}
## your code goes here
# Answered by Zuzanna 


# bar plot representing relationship between the gender and sound level preference
# creating th base of the plot with gender on the x axis and sound level preference on y axis
data %>% 
  ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  #adding first geom - the bar geom based on the mean of each gender
  geom_bar(
  stat = 'summary',
  fun = mean,
  width = .4
) + 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
  # adding custom colors to the plot
  scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))
#showing mean and standard deviation of the sound level preference based on gender
 data %>% 
   group_by(gender) %>%
   summarise(
    mean = mean(sound_level_pref, na.rm = TRUE),
    sd = sd(sound_level_pref, na.rm = TRUE)
    )

# box plot representing relationship between the gender and sound level preference
data %>% ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  #adding the boxplot geom
  geom_boxplot()
+ 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )
+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
  # adding custom colors to the plot
   scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))

# violin plot representing relationship between the gender and sound level preference
data %>% ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  # adding the violin geom
  geom_violin()
+ 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )
+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
# adding custom colors to the plot
   scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))

```

Explain your results in plain terms here (max 3 sentences):

There doesn't seem to be much difference in the sound level preference between people of different genders. As show in the bar plot it would seem that female participants seem to enjoy slightly higher levels of sound but as we can see in the violin plot the difference isn't actually that obvious. It also seems that there is not enough data to properly define the sound level likeness of people who identify as gender different that male or female.
---

#### Question 3

Shoe size could tell us something about general body size, which could also be connected to one's ability to hold your breath. In other words we predict that there is a positive relation between shoe size and how long time CogSci students can hold their breath. Try plotting the two sets of data against each other using a *scatter plot* (hint: both variables are continuous variables). You can make a scatter plot in `ggplot2` using the `geom_point()` function and plotting one variable on each axis. Use grouping in your plot to distinguish the relationship between shoe size and holding breath for the different gender subgroups, since we expect at least some of the groups to have different shoe sizes relative to each other. You can for instance use the `color` parameter within the `aes()` function to color by gender.

```{r question3}
## your code goes here
## Katarina 

ggplot(data = data)+
  geom_point(aes(x= shoesize, y = breathhold, color = gender))



```


Explain your results in plain terms here (max 3 sentences): 
Females breathhold was mostly similar (around 50 +-10sec) with all shoesizes. For male, the results were more differentiated but would not support the hypothesis that bigger shoesize means ability to hold breath for longer. The highest results were for female with shoesize around 38.
---

#### Question 4
# Answered by Alex

Is the `breathhold` variable normally distributed? Provide both visual (histogram and QQ-plot) and numeric (Shapiro-Wilk test and skewness/kurtosis values) support for your answer.

```{r question4}
## The variable 'breathhold' is NOT normally distributed, here's why...

## Quick Histogram
hist(data$breathhold)

## OR

## Custom Histogram
ggplot(data ,aes(x = breathhold)) +
  ggtitle('Histogram of the breathhold task') +
  labs(x = 'Length of Time', y = 'Density') +
  geom_histogram(aes(y = ..density..),
    colour = 'black', 
    fill = 'lightgreen', 
    bins = 50) +
  stat_function(fun = dnorm, 
                args = list(
                  mean = mean(data$breathhold, na.rm = TRUE),
                  sd = sd(data$breathhold, na.rm = TRUE)
                ), 
                colour= "red", size = 1) +
  theme_minimal() +
  xlim(range(data$breathhold))

## The above shows how the data is formed as a histogram and it can clearly be seen that the data does not follow a normal distribution curve, as there seems to almost be 2 peaks.

## Quick QQ-plot
qqnorm(data$breathhold)
qqline(data$breathhold)

## OR

## Custom QQ-Plot
ggplot(data, aes(sample = breathhold)) + 
  stat_qq() +
  stat_qq_line(colour = 'red') +
  labs(x = 'Theoretical Quantiles', y = 'Sample Quantiles') +
  ggtitle('Breathhold Q-Q Plot') +
  theme_bw()

## The above graph shows how the points do not follow the line of best fit (z-score) as although the values that lie within -1 and 1 of the theoretical quantiles seem to fit the tail end of the points start to deviate and become more spread out from one another. 

## Shapiro-Wilk test
shapiro.test(data$breathhold)
## Since the p-value of 0.000331 < 0.05 then we can assume the data is NOT normal and therefore significant

## Skewness/Kutosis values
round(pastecs::stat.desc(data$breathhold, basic = FALSE, norm = TRUE), digits = 2)
## Since both skew.2SE (1.88) and kurt.2SE (1.90) are both > 1 then the data is NOT normal and therefore significant

# In conclusion we can concur that the data is NOT normal and the variable 'breathhold' is not normally distributed.

```

---

#### Question 5

##Answered by Benjamin

Are the two balloon reaction time variables (`balloon` and `balloon_balance`) normally distributed? Provide visual (histogram and QQ-plot) and numeric (Shapiro-Wilk test and skewness/kurtosis values) support for your answer.

If they are not, then discuss your results below.

```{r question5, message=FALSE, warning=FALSE}

#Now the first step in deciding whether or not the data is normally distributed is a visual inspection. Here the histogram and QQ-plot will come in handy. 

#Creating a histogram for balloon_balance
ggplot(df, aes(balloon_balance)) +
  #using density as y-axis with 24 bins. 
  geom_histogram(aes(y = ..density..), color = "pink", fill = "green", bins = 24) + 
  #creating a line showing the expected shape of a normal distribution based on the mean and standard deviation of balloon_balance
  stat_function(fun = dnorm, args = list(mean = mean(df$balloon_balance, na.rm = TRUE), sd = sd(df$balloon_balance, na.rm = TRUE)), colour= "black", size = 1
                ) +
  theme_minimal()

#Using same approach for balloon_inflate
ggplot(df, aes(balloon_inflate)) +
  geom_histogram(aes(y = ..density..), color = "pink", fill = "green", bins = 24) + 
  stat_function(fun = dnorm, args = list(mean = mean(df$balloon_inflate, na.rm = TRUE), sd = sd(df$balloon_inflate, na.rm = TRUE)), colour= "black", size = 1
                ) +
  theme_minimal()

#Both density plots shows significant deviance from the expected bell-curve of a normal distribution visually indicating non-normality. 




#Making a qq-plot of balloon_balance and balloon_inflate with an identity line 

qqnorm(df$balloon_balance)
qqline(df$balloon_balance)

qqnorm(df$balloon_inflate)
qqline(df$balloon_inflate)

#Significant deviation from the identity lines representing normality in the qq-plots indicate non-normality


#Asking Shapiro Wilk if he can help us out with some testing of our variables - cbind() helps us do this jointly. 
round(pastecs::stat.desc(cbind(df$balloon_balance, df$balloon_inflate), basic = FALSE, norm = TRUE), digits = 3)

#Skewness and skew.2SE are greater than 1 for both variables indicating non-normality.
#Ballon balancing, W = 0.74, p < 0.05, and ballon inlating W = 0.74, p < 0.05 are both significantly non-normal since the p-value is less than 0.05. 

```

---

Explain your results in plain terms here (so grandma would understand): 
Hey Grandma! You see these bins here on the two graphs? - and the black line? Would you say they have roughly the same overall shape? Nahh i wouldnt either. What about these dots and the line? Yeaah ballon_inflation definitly fits for a while, but its still not a great match - Just like you and your first husband huhh grandma? Fits for a while and then.. arhh Haha im just kidding grandma. Yeaah i agree - sometimes you just have to look at the plain numbers... and he was way to old for you - i agree... And actually these fancy numbers here suggest the same thing for our variables - they just doesn't fit the normal curve very well... Yes yes calm down i know you are quite normal grandma - it was your old crazy husband that didn't fit... i agree.


That's all!
