---
title: "Portfolio exam - Part 1 | Methods 1 F23, CogSci @AU"
author: "YOUR NAME HERE, GROUP # HERE"
date: "12/10/2023"
output: html_document:
  toc: true
---

---

#### Deadline: October 12 2023 h23:59

#----------------------------------
# GENERAL INSTRUCTIONS
#----------------------------------

# 1. GITHUB:

# When you are done completing the below code, please push to Github with a commit comment (use _ for spaces in the message)

# You can push as many times as you would like. The last version that you commit before the assignment deadline is the one that will be reviewed for feedback.

# Please use your individual Github account to push assignments. Do not create a group account if you are working in a group.

#----------------------------------

# 2. PUSH CODE & KNIT HTML MARKDOWN DOCUMENT.

# Please push your code AND a knit HTML file of the document to Github. 

#----------------------------------

# 3. GROUPWORK

**Please write your name in the `author` field above, followed by your Group #.**

# If you are working in groups, it should be possible to grade your code AND your written text individually. 

# Therefore, please indicate your group number at the top of your markdown file and indicate your group members' names below.

#----------------------------------

# 4. Suppressing code output.

# Please suppress code output that is not for evaluation (e.g., printouts of data frames, variable values, etc.) by setting echo = FALSE in chunks where only code is being run. (Look in Class3A_PHB_Rmd from class 3 for more info on echo = FALSe and also further down in this document). This will make it much easier to evaluate your HTML output files.

#----------------------------------

#### Group members' names.

# Please indicate your group members' names here.
My group members' names are: []

#### Introduction

The goal of this exam is to write a short data mining report on the CogSci Intro Week Personality Test Data in which you answer the following questions **in prose, code and graphs**.

First of all, let's start by looking at the setup chunk. If you need to load packages or set your working directory, do so here:

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)
## your code goes here
```

Now you have to import the personality data from the 'cogsci_personality_2023.csv' file; i.e., the one we have worked with in both class 2 and class 3. Note the filename is in lowercase and it is located in the /renv_cogsci_e23 folder on UCloud. 

Once you have imported it, use the `head()` function to print the first 10 lines of the data set.

```{r load_data}
## your code goes here

```

Once you are done loading the data, you can start working on the questions below.

### Question 1

#### Question 1.1

Who can hold their breath the longest on average — those with right or left ocular dominance? Notice that the column is called `ocular_dom`, and that right ocular dominance is indicated in the column with 'Right', while left ocular dominance is indicated in the column with 'Left'. Therefore, you want to only filter out the data in this column which corresponds to either "Right" or "Left".

Plot the data using `ggplot2` to find out the answer to your question. The plots should include error bars depicting the standard error of the mean: you can add these using the `geom_errorbar()` function and specifying `stat = "summary", fun.data = "mean_se"`. 

#### Question 1.2

Then use the `mean()` and `sd()` functions *within a tidyverse pipe* to make a summary data set, in which you show mean and standard deviation of the two eye dominance groups.

Bonus question: If you feel brave, you can instead try making a boxplot (`geom_boxplot()`) or a violin plot (`geom_violin()`) which are better at representing the actual distribution of the data (compared to a bar plot, which only depicts mean and standard deviation).


```{r question1}
## your code goes here

```

Explain your results in plain terms here (max 3 sentences): 

---

#### Question 2

Who likes silence vs. noise best -- by gender? Also in this case you should plot the data using `ggplot2` (including error bars depicting the standard error of the mean), then use the `mean()` and `sd()` functions to find mean and standard deviation of the two genders (still making a summary data set with tidyverse and pipes).

Bonus question: If you feel brave, you can instead try making a boxplot (`geom_boxplot()`) or a violin plot (`geom_violin()`) which are better at representing the actual distribution of the data (compared to a bar plot, which only depicts mean and standard deviation).

```{r question2}
## your code goes here


# bar plot representing relationship between the gender and sound level preference
# creating th base of the plot with gender on the x axis and sound level preference on y axis
data %>% 
  ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  #adding first geom - the bar geom based on the mean of each gender
  geom_bar(
  stat = 'summary',
  fun = mean,
  width = .4
) + 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
  # adding custom colors to the plot
  scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))
#showing mean and standard deviation of the sound level preference based on gender
 data %>% 
   group_by(gender) %>%
   summarise(
    mean = mean(sound_level_pref, na.rm = TRUE),
    sd = sd(sound_level_pref, na.rm = TRUE)
    )

# box plot representing relationship between the gender and sound level preference
data %>% ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  #adding the boxplot geom
  geom_boxplot()
+ 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )
+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
  # adding custom colors to the plot
   scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))

# violin plot representing relationship between the gender and sound level preference
data %>% ggplot(aes(x = gender, y = sound_level_pref, fill = gender)) +
  # adding the violin geom
  geom_violin()
+ 
  #adding a second geom - the error bar showing the standard error 
  geom_errorbar(
    stat = 'summary',
    fun.data = mean_se,
    width = .2
  )
+
  # labeling the plot to make it more legible
  labs(x = "Gender", 
       y = "Sound level preference", 
       title = "Relatonship between the gender and sound level preference",
       subtitle = "",
       caption = "Data source: CogSci Personality Test 2023") +
# adding custom colors to the plot
   scale_fill_manual(values = c(
    '#D7FCD4',
    '#854798',
    '#EDA2C0'
  ))

```

Explain your results in plain terms here (max 3 sentences):

There doesn't seem to be much difference in the sound level preference between people of different genders. As show in the bar plot it would seem that female participants seem to enjoy slightly higher levels of sound but as we can see in the violin plot the difference isn't actually that obvious. It also seems that there is not enough data to properly define the sound level likeness of people who identify as gender different that male or female.
---

#### Question 3

Shoe size could tell us something about general body size, which could also be connected to one's ability to hold your breath. In other words we predict that there is a positive relation between shoe size and how long time CogSci students can hold their breath. Try plotting the two sets of data against each other using a *scatter plot* (hint: both variables are continuous variables). You can make a scatter plot in `ggplot2` using the `geom_point()` function and plotting one variable on each axis. Use grouping in your plot to distinguish the relationship between shoe size and holding breath for the different gender subgroups, since we expect at least some of the groups to have different shoe sizes relative to each other. You can for instance use the `color` parameter within the `aes()` function to color by gender.

```{r question3}
## your code goes here

ggplot(data = data)+
  geom_point(aes(x= shoesize, y = breathhold, color = gender))



```


Explain your results in plain terms here (max 3 sentences): 
Females breathhold was mostly similar (around 50 +-10sec) with all shoesizes. For male, the results were more differentiated but would not support the hypothesis that bigger shoesize means ability to hold breath for longer. The highest results were for female with shoesize around 38.
---

#### Question 4

Is the `breathhold` variable normally distributed? Provide both visual (histogram and QQ-plot) and numeric (Shapiro-Wilk test and skewness/kurtosis values) support for your answer.

```{r question4}
## your code goes here

```

---

#### Question 5

Are the two balloon reaction time variables (`balloon` and `balloon_balance`) normally distributed? Provide visual (histogram and QQ-plot) and numeric (Shapiro-Wilk test and skewness/kurtosis values) support for your answer.

If they are not, then discuss your results below.

```{r question5}
## your code goes here

```

---

Explain your results in plain terms here (max 3 sentences): 


That's all!
